% !Mode\dots ``TeX:UTF-8''

\section{Algorithm based on  continued fraction}
\label{sec:contalg}
In this section, we first recall {\em Descartes' rule of signs}, which gives a bound on the number of positive real roots. Then the {\em Vincent theorem}, which  ensures
the termination of algorithms based on  continued fractions, is presented. Finally, we review an algorithm of real root isolation based on  continued fractions.


As usual, $\deg(p)$ denotes the degree of univariate polynomial $p$. The derivative of polynomial $p$ with respect to the only variable is denoted by $p'$ and $\gcd(f,g)$ means the greatest common divisor of polynomials $f$ and $g$.
\rev{
\begin{example}
	\label{ex:exp1}
	Consider the polynomial
	\[p_1(x)=x^6+2x^5-4x^4+x^3+10x^2-5x+5.\]
	\[\deg(p_1)=6, p_1'=6x^5+10x^4-16x^3+3x^2+20x-5.\]
\end{example}
}

\begin{note}[Sign variation]
Let $S=\left\{ a_0,a_1,\ldots,a_n \right\}$ be a finite sequence of non-zero real numbers. Define $V(S)$, the {\em sign variation} of $S$, as follows.
\[V(S)=0\ \text{ if } |S|\le1,\]
\[  V(a_0,\ldots,a_{n-1},a_n)=  \left\{\begin{aligned}
 &  V(a_0,\ldots,a_{n-1})+1 \text{ if }a_{n-1}a_n<0;\\
&V(a_0,\ldots,a_{n-1}), \text{ otherwise}.\\
	\end{aligned}
	\right.
\]
If some elements of $S$ are zero, remove those zero-elements to get a new sequence and define $V(S)$ to be the sign variation of this new sequence.

\end{note}



%\rev{There are many studies about  real roots for a given polynomial,  and {\em Descartes' rule of signs} is one of beautiful results among them.  It  provides an upper bound of the number of positive real roots.}


\begin{theorem}[Descartes' rule of signs] \label{thm:des}
  Suppose $p=\sum_{i=0}^na_ix^i\in\RR[x]$ has $m$ positive real roots, counted with multiplicity. Set $V(p)=V(a_0,a_1,\ldots,a_n)$. Then $m\le V(p)$, and $V(p)-m$ is even.
\end{theorem}

%\rev{ For {\em Example \ref{ex:exp1}}, $V(p_1)=V(5,-5,10,1,-4,2,1)=4$.  {\em Descartes' rule of signs} is an useful tool to estimate the number of  real roots. And if $V(p)=1$, then employ this rule, there is exact one  positive root.  But when $V(p)>1$, the rule can not directly isolate the real roots. {\em Vincent's theorem} in the following will overcome this restriction. {\em Vincent's theorem} says that if we give a small enough interval, then it can map this interval to entire positive domain and $V(p)\le 1$.}



\begin{theorem}[Vincent's theorem]\label{thm:vin}
  Let $p(x)$ be a real polynomial of degree $n$ which has only simple roots. It is possible to determine a positive quantity $\delta$ so that for every pair of positive real numbers $a$ and $b$ with $|b-a| < \delta$, the coefficients sequence of every transformed polynomial of the form
  $  p(x) = (1+x)^{n}p(\frac{a+bx}{1+x}) $
		  has exactly 0 or 1 sign variation. The second case is possible if and only if $p(x)$ has a single root within $(a,b)$.
\end{theorem}

\begin{definition}  We define the following transformations for a univariate polynomial $p(x)=a_nx^n+a_{n-1}x^{n-1}+\cdots+a_1x+a_0,n>0$.
  \begin{eqnarray*}
  lc(p(x))&=&a_n,\\
  R(p(x))&=&x^n(p(\frac{1}{x})),\\
  H_\lambda(p(x))&=&p(\lambda x),\\
  T(p(x))&=&p(x+1), \\
  D(p(x))&=&a_nx^{n-1}+a_{n-1}x^{n-2}+\cdots+a_2x+a_1.
  \end{eqnarray*}
\end{definition}

$T(p)$  is also called	{\em Taylor shift one} \cite{ger04,joh05}.

\begin{algorithm}
\SetAlgoCaptionSeparator{.}
\caption{\algm \label{alg:main}}
\DontPrintSemicolon
\KwIn{ A non-zero polynomial $p(x)\in \ZZ[x] $. }
\KwOut{$I$,  a set of real root isolating intervals of $p(x)$. }
$I=\emptyset$; \;
\If {$\deg(p)$  equals to $0$}  {return $I$;}
$p=\frac{p}{\gcd(p,p')}$; \tcc{ square free}
\If {$p(0)$   equals to  $0$ } { $I.add([0,0])$; \tcc{ add $[0,0]$ to set $I$}
$p=D(p)$; \tcc{derivative of $p$ }
}
$I$.addAll(\algcf($p$)); \;
\tcc { add all the positive root intervals to set $I$ } \tcc{ \algcf\ is described as {\em Algorithm \ref{alg:cf}}}
$p=p(-x)$;\;
$I$.addAll(-\algcf($p$));
\end{algorithm}


%In our experiments when {\em Algorithm \ref{alg:up}} is used for computing upper bounds, $T(p)$  takes  more than ninety percent of running time\footnote{The result of  GNU gprof.}.  We have considered methods in \cite{ger04} for computing $T(p)$, but   finally we chose the  classical method (Horner's method) for its simplicity. In future work we will use Divide \& Conquer method which is the fastest in \cite{ger04}. We think \rev{further} substituting will still improve the performance of our method.
{\em Vincent's theorem} provides  a possible method to isolate the real roots. But it needs to divide $\RR$ into  many small intervals and the width of these intervals is smaller than a given  finite value. So {\em Vincent's theorem} cannot be used to isolate the real roots directly since $\RR$ is infinite. For employing  {\em Vincent's theorem}, we need to discard the intervals containing no real roots. {\em Algorithm \ref{alg:lb}} provides a lower  bound $lb$ of positive roots for given $p$.  Since the interval  $(0,lb)$ contains no real roots,  we can safely  discard $(0,lb)$ when isolate the real roots.


\begin{algorithm}
\SetAlgoCaptionSeparator{.}
\caption{\lb \label{alg:lb}}
\DontPrintSemicolon
\KwIn{ $p\in\ZZ[x],\ lc(p)\neq 0 $. }
\KwOut{$root\_lb$, a lower bound of positive roots of $p$. }
$p=R(p)$;\;
\lIf{$lc(p)<0$}{	$p=-p$;}
$root\_lb=$\up($p$); \tcc{\up\ is described as {\em Algorithm \ref{alg:up}}}
\end{algorithm}


\begin{definition}
$  intvl(a,b,c,d)=  \left\{\begin{aligned}
&  (\min\left\{ \frac{a}{c},\frac{b}{d} \right\},\ \max\left\{ \frac{a}{c},\frac{b}{d} \right\} ) &\text{ if } cd\neq0;\\
& (0,\infty), &\text{ otherwise}.\\
	\end{aligned}
	\right.
$
\end{definition}


Using the above notations and definitions, an algorithm for isolating all the real roots of a nonzero univariate polynomial is described as {\em Algorithm \ref{alg:main}}.
{\em Algorithm \ref{alg:cf}}, which is a slight modification of the algorithm in \cite{akr08}, is presented here to make our subsequent description clearer.

Continued fractions based procedures will continue subdividing the considered interval into two subintervals and make a one to one map from $(a,b)$ to $(0,+\infty)$ by $  p(x) = (1+x)^{n}p(\frac{a+bx}{1+x})$ until $V(p)$ equals $1$ or $0$. Informally, {\em Algorithm \ref{alg:main}} is employing the above map to magnify the considered interval  to $(0,+\infty)$.  Through {\em Descartes' rule of signs}, there is no positive real roots if  $V(p)=0$. In this case we delete the interval from the considered interval set.   When $V(p)=1$, this interval contains exact one real root by {\em Descartes' rule of signs}. In this case we throw away the interval from the considered interval set. When $V(p)>1$, since it cannot be determined how many real roots are contained in this interval, we divide the considered interval into two subintervals. In this case we throw away the interval from the considered interval set and add two subintervals to the considered interval set. Repeating this procedure, we can divide the interval into  subintervals until  there is no width of  the corresponding original interval greater or equal to $\delta$ in considered interval set.   Then there is at most one real root in every interval in the considered interval set and then the procedure will terminate. This is also
the reason why {\em Theorem \ref{thm:vin}}  can guarantee the termination of {\em Algorithm \ref{alg:main}}.



\begin{algorithm}
	\SetAlgoCaptionSeparator{.}
	\caption{\algcf \label{alg:cf}}
	\DontPrintSemicolon
	\KwIn{ A squarefree polynomial $p \in \ZZ[x] \setminus \{0\}$. }
	\KwOut{ $roots$, a list of isolating intervals of positive roots of $p$. }
	
	$roots=\emptyset$;
	$s=V(p)$;\;

	$intstack=\emptyset$;
	$intstack$.add($\{1,0,0,1,p,s\}$);\;
	\While{$intstack\neq \emptyset$} {
		$\{a,b,c,d,p,s\}=intstack.$pop();\tcc{pop  the first element}
		$\lambda=\lb(p)$;\;
		\lIf{$\lambda \ge 1$ \label{log:branch}}{
			$\{ a,c,p \}=\{ \lambda  a,\lambda c ,H_\lambda(p) \}$;
			$\{ b,d,p \}=\{   a+b,c+d ,T(p) \}$;\;
			\lIf { $p(0)= = 0$}{
				$roots$.add$([\frac{b}{d},\frac{b}{d}])$ ;
				$p=\frac{p}{x}$;}
			$s=V(p)$;
			\lIf {$s= = 0$}{
				continue;
			}
			\lElseIf{$s= = 1$}{		
				$roots$.add($intvl(a,b,c,d)$);
				continue;
			}
		}
		$ \left\{ p_1,a_1,b_1,c_1,d_1,r \right\}=\left\{ T(p),a,a+b,c,c+d,0 \right\}$
		
		\lIf{$p_1(0)= =0$}{
			$roots$.add($[\frac{b_1}{d_1},\frac{b_1}{d_1}]$);
			$p_1=\frac{p_1}{x};r=1$;
		}
		$s_1=V(p_1)$;
		$\left\{ s_2,a_2,b_2,c_2,d_2 \right\}=\left\{ s-s_1-r,b,a+b,d,c+d \right\}$;
		%$ s_2 = s- s_1 - r; a_2 = b; b_2 = a + b; c_2 = d; d_2 = c + d$;\;
		\lIf {$s_2>1$ }{
			$ p_2= (x+1)^{\deg(p)}T(p)$;\;
			\lIf {$p_2(0)= = 0$}{
				$p_2=\frac{p_2}{x}$;
				$s_2=V(p_2)$;
			}
		}
		\lIf{$s_1= = 1$ }{
			$roots$.add($intvl(a_1,b_1,c_1,d_1)$);\;
		}
		\lElseIf{$s_1>1$ }{
			$intstack$.add($\{a_1,b_1,c_1,d_1,p_1,s_1\}$);\;
		}
		
		\lIf{$s_2= = 1$}{
			$roots$.add($intvl(a_2,b_2,c_2,d_2)$);\;
		}
		\lElseIf{$s_2>1$ }{
			$intstack$.add($\{a_2,b_2,c_2,d_2,p_2,s_2\}$);\;
		}
	}
\end{algorithm}

